title: "HM Index Teen Validation Study 1"
author: "masked for review"
date: "3/2/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a script to conduct analysis for validation of the adolescent Healthy Minds (HM) Index data that was collected by the Character Lab: <https://characterlab.org/about/>. Data can be found on the HM Index project page on the (masked for review) Open Science Framework: https://osf.io/aw7bz/?view_only=954a8b1dea9e45a9b52a0194e726d106

Load the necessary libraries.

```{r libraries}

library(renv)
library(here)
library(readr)
library(tidyverse)
library(janitor)
library(psych)
library(ggplot2)
library(assertr)
library(lmSupport)
library(irr)

```

## Import data


```{r import}

hmindex_charlab_017 <- read.csv("/pathToData/HMINDEXRETEST_F20_A_0929.csv")
hmidx17retest <- read.csv("/pathToData/TeenStudy1_data_HMINDEXRETEST_W21_Followup_A_share.csv")

```


## Clean and convert the raw data

``` {r cleanRaw, echo=FALSE}

# Inspect content of HM Index items that need renaming
hmindex_charlab_017[1:3,23:39] 
hmidx17retest[1:3,22:38] 

# As expected, continue with renaming for ease of analysis
newnames <- c("purpose6", "purpose9", "purpose12", "purpose14",
              "awareness1", "awareness8", "awareness12", "awareness13",
              "insight4", "insight5", "insight14",
              "connection1", "connection17", "connection21",
              "connection2", "connection3", "connection11")
names(hmindex_charlab_017)[23:39] <- newnames
names(hmidx17retest)[22:38] <- newnames

# Same for demographic variables
hmindex_charlab_017[1:3,138:144]
hmidx17retest[1:3,137:143]
newnames <- c("age", "gender", "race", "region",
              "urbanicity", "ethnicity", "engfluency")
names(hmindex_charlab_017)[138:144] <- newnames
names(hmidx17retest)[137:143] <- newnames

# Second row has all the item wordings
item_content <- hmindex_charlab_017[1, 23:39]
item_content <- (t(item_content))
item_content <- as.data.frame(item_content)
item_content <- tibble::rownames_to_column(item_content)
# write_csv(item_content, "item_content_charlab17hmidx.csv")

item_content <- hmidx17retest[1, 22:38]
item_content <- (t(item_content))
item_content <- as.data.frame(item_content)
item_content <- tibble::rownames_to_column(item_content)
# write_csv(item_content, "item_content_charlab17hmidx_retest.csv")

# First two rows are duplicates of column names; remove
hmindex_charlab_017 <- hmindex_charlab_017[3:length(hmindex_charlab_017$StartDate), ]
hmidx17retest <- hmidx17retest[3:length(hmidx17retest$StartDate), ]

# Re-write and re-import the data
# write.csv(hmindex_charlab_017, "/pathToData/hmindex_charlab_017_2March2023.csv")
hmindex_charlab_017 <- read.csv("/pathToData/clean/hmindex_charlab_017_07Sep21_num_14July2022.csv")
# write.csv(hmidx17retest, "/pathToData/hmidx17retest.csv")
hmidx17retest <- read.csv("/pathToData/data/clean/hmidx17retest.csv")

# Convert blanks to NA
hmindex_charlab_017[hmindex_charlab_017 == ""] <- NA
hmidx17retest[hmidx17retest == ""] <- NA

# Convert response duration
hmindex_charlab_017 <- hmindex_charlab_017 %>% rename(duration = 7)
hmindex_charlab_017$duration <- as.character(hmindex_charlab_017$duration)
hmindex_charlab_017$duration <- as.numeric(hmindex_charlab_017$duration)

hmidx17retest <- hmidx17retest %>% rename(duration = 7)
hmidx17retest$duration <- as.character(hmidx17retest$duration)
hmidx17retest$duration <- as.numeric(hmidx17retest$duration)

# make sure all consent = yes for initial session
table(hmindex_charlab_017$Q1965) 

# only use where consent = yes
hmidx_consented <- hmindex_charlab_017[hmindex_charlab_017$Q1965 == 1,]
hmidx17retest_consented=subset(hmidx17retest, AnonymizedStudentID %in% hmidx_consented$AnonymizedStudentID)
table(hmidx_consented$Q1965) # check

# Check whether numeric imported as numeric
typeof(hmidx_consented$insight14) # imported as integer
typeof(hmidx_consented$Q2000) # imported as integer
typeof(hmidx17retest_consented$insight14) # imported as character

# Convert to numeric, check total number for each response category.
# Confirmed that table of tallies of responses before and after are identical.
beforeconvert <- lapply(hmidx_consented[24:138], table)
hmidx_consented[24:138] <- lapply(hmidx_consented[24:138], as.character)
# Must convert to character before numeric; converting directly from
# integer to numeric recodes values.
hmidx_consented[24:138] <- lapply(hmidx_consented[24:138], as.numeric)
afterconvert <- lapply(hmidx_consented[24:138], table)
identical(beforeconvert, afterconvert)

hmidx_consented$race_wnw <- car::recode(hmidx_consented$race, 
                                            "7 = 1; 1 = 2; 2 = 2; 3 = 2; 4 = 2; 5 = 2; 6 = 2; 8 = 2")

hmidx_consented$race_wnw <- car::recode(hmidx_consented$race_wnw, 
                                            "1 = 'white'; 2 = 'not white'")

hmidx_consented$gender <- car::recode(hmidx_consented$gender, 
                                          "1 = 'male'; 2 = 'female'; 3 = 'nonbinary'; 4 = 'other'")

hmidx_consented$region <- car::recode(hmidx_consented$region, 
                                          "1 = 'midwest'; 2 = 'northeast'; 3 = 'south'; 4 = 'West'; 5 = 'PR or other'")

# Now for retest
# beforeconvert <- lapply(hmidx17retest_consented[23:137], table)
# hmidx17retest_consented[23:137] <- lapply(hmidx17retest_consented[23:137], as.character)
# hmidx17retest_consented[23:137] <- lapply(hmidx17retest_consented[23:137], function(x) as.integer(factor(x)))
# hmidx17retest_consented[23:137] <- lapply(hmidx17retest_consented[23:137], as.character)
# # must convert to character before numeric; converting directly from integer to numeric recodes values.
# hmidx17retest_consented[23:137] <- lapply(hmidx17retest_consented[23:137], as.numeric)
# afterconvert <- lapply(hmidx17retest_consented[23:137], table)
# identical(beforeconvert, afterconvert)

hmidx_consented$gender <- as.factor(hmidx_consented$gender)
hmidx_consented$race <- as.factor(hmidx_consented$race)
hmidx_consented$region <- as.factor(hmidx_consented$region)
hmidx_consented$race_wnw <- as.factor(hmidx_consented$race_wnw)

# Create subsets to simplify retest analysis
vars1=c('AnonymizedStudentID',"gender", "race", "region",
       "urbanicity", "ethnicity", "engfluency","age","purpose6", "purpose9", "purpose12", "purpose14",
       "awareness1", "awareness8", "awareness12", "awareness13",
       "insight4", "insight5", "insight14",
       "connection1", "connection17", "connection21",
       "connection2", "connection3", "connection11",
       'duration')

vars2=c('AnonymizedStudentID',"purpose6", "purpose9", "purpose12", "purpose14",
        "awareness1", "awareness8", "awareness12", "awareness13",
        "insight4", "insight5", "insight14",
        "connection1", "connection17", "connection21",
        "connection2", "connection3", "connection11",
        'duration')

hmidx_test_subset=hmidx_consented[vars1]
hmidx_retest_subset=hmidx17retest_consented[vars2]

data=merge(hmidx_test_subset,hmidx_retest_subset,by="AnonymizedStudentID",all.x=T,all.y=T)

# convert to minutes for interpretability
data$minutes.x <- data$duration.x / 60
data$minutes.y <- data$duration.y / 60

```


## Subset & merge

Code is silenced for this chunk.

``` {r merge, echo=FALSE}

# PRE variables
vars1=c('AnonymizedStudentID',"gender", "race", "region",
       "urbanicity", "ethnicity", "engfluency","age","purpose6", "purpose9",
       "purpose12", "purpose14",
       "awareness1", "awareness8", "awareness12", "awareness13",
       "insight4", "insight5", "insight14",
       "connection1", "connection17", "connection21",
       "connection2", "connection3", "connection11",
       'duration','hmindex','connection','purpose','insight','awareness')

# POST variables
vars2=c('AnonymizedStudentID',"purpose6", "purpose9", "purpose12", "purpose14", "awareness1", "awareness8", "awareness12", "awareness13",
        "insight4", "insight5", "insight14",
        "connection1", "connection17", "connection21",
        "connection2", "connection3", "connection11",
        'duration','hmindex','connection','purpose','insight','awareness')

# Subset
hmidx_test_subset=hmindex_charlab_017[vars1]
hmidx_retest_subset=hmidx17retest[vars2]

# Merge
data=merge(hmidx_test_subset,hmidx_retest_subset,by="AnonymizedStudentID",all.x=T,all.y=T)
data=data[!is.na(data$AnonymizedStudentID),]

# convert duration to minutes
data$minutes.x <- data$duration.x / 60
data$minutes.y <- data$duration.y / 60

```

## Process outcomes

```{r processData}

hmindex_charlab_017=hmidx_consented

# WHO-5 Well-being Index
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  rowwise() %>%
  dplyr::mutate(who = sum(Q2089,
                          Q2090,
                          Q2091,
                          Q2092,
                          Q2093,
                          na.rm = TRUE
  ))

psych::alpha(hmindex_charlab_017 %>%
               dplyr::select(Q2089,
                             Q2090,
                             Q2091,
                             Q2092,
                             Q2093))


# EPOCH: connectedness
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    epconnect = (Q2107 + Q2108 + Q2109 + Q2110) / 4,
    na.rm = TRUE
  )

psych::alpha(hmindex_charlab_017 %>%
               dplyr::select(
                 Q2107, Q2108, Q2109, Q2110
               ))


# Comprehensive Inventory of Thriving Loneliness subscale
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    lonely = (Q2112 + Q2113 + Q2114) / 3,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017 %>%
               dplyr::select(Q2112, Q2113, Q2114))


# Diener Satisfaction with Life Scale-Children (SWLS-C)
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  rowwise() %>%
  dplyr::mutate(satwlife = sum(Q2116,
                               Q2117,
                               Q2118,
                               Q2119,
                               Q2120,
                               na.rm = TRUE
  ))

psych::alpha(hmindex_charlab_017 %>%
               dplyr::select(
                 Q2116,
                 Q2117,
                 Q2118,
                 Q2119,
                 Q2120
               ))


# francis ----
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    francis_purpose = Q2077,
    na.rm = TRUE
  )


# Kessler Psychological Distress Scale (K10)
# Questions 3 and 6 are not asked if the preceding question was ‘none of the time’ in which case questions 3 and 6 would automatically receive a score of one.
hmindex_charlab_017$Q2098[hmindex_charlab_017$Q2097 == 1] <- 1
hmindex_charlab_017$Q2099[hmindex_charlab_017$Q2100 == 1] <- 1
# raw total of psydistress1-psydistress10  =

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    psydistress = Q2096 + Q2097 + Q2098 +
      Q2099 + Q2100 + Q2101 +
      Q2103 + Q2104 + Q2105,
    na.rm = TRUE
  )

psych::alpha(hmindex_charlab_017 %>%
               dplyr::select(
                 Q2096,Q2097,Q2098,
      Q2099,Q2100, Q2101,
      Q2103,Q2104,Q2105
               ))

# scale computation: insight ----
# insight: entire scale

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    insight = (insight4 + insight5 + insight14) / 3,
    na.rm = TRUE
  )

insight_items <- hmindex_charlab_017 %>%
  dplyr::select(
    insight4, insight5, insight14
  )

# scale computation: insight convergent ----

# DERS Limited access to emotion regulation strategies

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(ders_lack_emoreg = (Q2064 + Q2065 + Q2068 + Q2067) / 4, na.rm = TRUE)

psych::alpha(hmindex_charlab_017 %>% dplyr::select(Q2064, Q2065, Q2068, Q2067))

# DERS Non-acceptance of emotional responses

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    ders_nonaccept = (Q2066 + Q2067 + Q2069) / 3,
    na.rm = TRUE
  )

psych::alpha(hmindex_charlab_017 %>% dplyr::select(Q2066, Q2067, Q2069))


# reappraisal

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    reappraisal = (Q2057 + Q2058 + Q2059 + Q2060 + Q2061 + Q2062) / 6,
    na.rm = TRUE
  )

psych::alpha(hmindex_charlab_017 %>% 
               dplyr::select(
                 Q2057, Q2058, Q2059, Q2060, Q2061, Q2062
                 ))




# scale computation: awareness ----
# Awareness: entire scale
# Calculate removing any Ps who failed to respond to any items

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    awareness = (awareness1 +  awareness8
                 + awareness12 + awareness13) / 4,
    na.rm = TRUE
  )

awareness_items <- hmindex_charlab_017 %>%
  dplyr::select(
    awareness1,  awareness8,
    awareness12, awareness13
  )

# scale computation: awareness convergent ----

# relativity of thoughts
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    chimea_rel = (Q2053 + Q2054 + Q2055) / 3,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017 %>% 
               select(Q2053, Q2054, Q2055))


# CHIME-A
# Awareness of internal experiences
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    chimea_awa = (Q2004 + Q2005 + Q2006) / 3,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017 %>% 
               select(Q2004, Q2005, Q2006)) # .82

# Decentering and nonreactivity
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    chimea_dec = (Q2013 + Q2014 + Q2015) / 3,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017 %>% 
               select(Q2013, Q2014, Q2015))


# Acting with awareness
hmindex_charlab_017$chimea_act_1r <- 8 - hmindex_charlab_017$Q2000
hmindex_charlab_017$chimea_act_2r <- 8 - hmindex_charlab_017$Q2001
hmindex_charlab_017$chimea_act_3r <- 8 - hmindex_charlab_017$Q2002

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    chimea_act = (chimea_act_1r +
                    chimea_act_2r + chimea_act_3r) / 3,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017 %>% select(chimea_act_1r, chimea_act_2r, chimea_act_3r))


# ESQ Attention
# Reverse code esq_attn3 and esq_attn4.
hmindex_charlab_017$esq_attn3r <- 8 - hmindex_charlab_017$Q2010
hmindex_charlab_017$esq_attn4r <- 8 - hmindex_charlab_017$Q2011

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    esq_attn = (Q2008 + Q2009 +
                  esq_attn3r +
                  esq_attn4r) / 4,
    na.rm = TRUE
  )

psych::alpha(hmindex_charlab_017
             %>% dplyr::select(
               Q2008,
               Q2009, esq_attn3r, esq_attn4r
             )) 

# The Mindful Attention Awareness Scale (MAAS-15)
# we did not use maas12
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    maas = (Q2017 + Q2018 + Q2019 +
              Q2020 + Q2021 + Q2022 +
              Q2023 + Q2024 + Q2025 +
              Q2026 + Q2027 + Q2028 +
              Q2029 + Q2030) / 14,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017
             %>% dplyr::select(
               Q2017, Q2018, Q2019,
                 Q2020, Q2021, Q2022,
                 Q2023, Q2024, Q2025,
                 Q2026, Q2027, Q2028,
                 Q2029, Q2030
             )) #

# scale computation: connection ----

# Connection: entire scale
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    connection = (connection1 + connection2 +
                    connection3
                  + connection11 + connection17
                  + connection21) / 6,
    na.rm = TRUE
  )

connection_items <- hmindex_charlab_017 %>%
  dplyr::select(
    connection1, connection2, connection3,
    connection11,  connection17,
    connection21
  )

psych::alpha(connection_items)


# scale computation: connection convergent  ----

# positive relations
hmindex_charlab_017$posrels_2r <- 6 - hmindex_charlab_017$Q2046
hmindex_charlab_017$posrels_3r <- 6 - hmindex_charlab_017$Q2047
hmindex_charlab_017$posrels_6r <- 6 - hmindex_charlab_017$Q2050

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  rowwise() %>%
  dplyr::mutate(posrels = sum(Q2045,
                              posrels_2r, posrels_3r, Q2048, Q2049,
                              posrels_6r, Q2051,
                              na.rm = TRUE
  ))

psych::alpha(hmindex_charlab_017 %>% dplyr::select(
  Q2045,
  posrels_2r, posrels_3r, Q2048, Q2049,
  posrels_6r, Q2051
))

# compassion
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    compassion = (Q2032 + Q2033 +
                    Q2034 + Q2035 + Q2036) / 5,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017 %>% dplyr::select(
  Q2032, Q2033, Q2034, Q2035, Q2036
))

# general trust
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    gentrust = (Q2038 +
                  Q2039 + Q2040 + Q2041 +
                  Q2042 + Q2043) / 6,
    na.rm = TRUE
  )
psych::alpha(hmindex_charlab_017 %>% dplyr::select(
  Q2038,
    Q2039, Q2040, Q2041,
    Q2042, Q2043
))


# scale computation: purpose ----

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    purpose = (
               purpose6
               + purpose9 
               + purpose12 + purpose14) / 4,
    na.rm = TRUE
  )

purpose_items <- hmindex_charlab_017 %>%
  dplyr::select(
    purpose6,
    purpose9,
    purpose12,
    purpose14
  )

# scale computation: purpose convergent ----

# costin purpose
hmindex_charlab_017$costin_purpose_3r <- 8 - hmindex_charlab_017$Q2074
hmindex_charlab_017$costin_purpose_4r <- 8 - hmindex_charlab_017$Q2075

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    costin_purpose = (Q2072 + Q2073 +
                        costin_purpose_3r + costin_purpose_4r) / 4,
    na.rm = TRUE
  )

psych::alpha(hmindex_charlab_017 %>% dplyr::select(
  Q2072, Q2073,
  costin_purpose_3r, costin_purpose_4r
))

# meaning in life
hmindex_charlab_017$meaning_5r <- 8 - hmindex_charlab_017$Q2083

hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    meaning = (Q2079 + Q2080 +
                 Q2081 + Q2081 + meaning_5r) / 5,
    na.rm = TRUE
  )

psych::alpha(hmindex_charlab_017 %>% dplyr::select(
  Q2079, Q2080,
    Q2081, Q2081, meaning_5r
))

# whole hm index -----
hmindex_charlab_017 <- hmindex_charlab_017 %>%
  dplyr::mutate(
    hmindex = (awareness + connection +
                        purpose + insight) / 4,
    na.rm = TRUE
  )


```
## Factor analysis

```{r factors}

# FA on all items  -----

parallel_fa_all_items <- fa.parallel(hmindex_charlab_017 %>%
  dplyr::select("purpose6", "purpose9", "purpose12", "purpose14",
                "awareness1", "awareness8", "awareness12", "awareness13",
                "insight4", "insight5", "insight14",
                "connection1", "connection17", "connection21",
                "connection2", "connection3", "connection11"))

parallel_fa_all_items

print(evalues <- eigen(polychoric(hmindex_charlab_017 %>%
  dplyr::select("purpose6", "purpose9", "purpose12", "purpose14",
                "awareness1", "awareness8", "awareness12", "awareness13",
                "insight4", "insight5", "insight14",
                "connection1", "connection17", "connection21",
                "connection2", "connection3", "connection11"))$rho)$values)
# 4 eigenvalues > 1.0

fa_vss <- vss(hmindex_charlab_017 %>%
  dplyr::select("purpose6", "purpose9", "purpose12", "purpose14",
                "awareness1", "awareness8", "awareness12", "awareness13",
                "insight4", "insight5", "insight14",
                "connection1", "connection17", "connection21",
                "connection2", "connection3", "connection11"),
n = 9,
rotate = "oblimin", 
diagonal = FALSE,
fm = "minres",
cor = "poly" 
) 

summary(fa_vss)
# 1 and 3 factors supported

fa_3factor <- psych::fa(hmindex_charlab_017 %>%
                          dplyr::select("purpose6", "purpose9", "purpose12", "purpose14",
                                        "awareness1", "awareness8", "awareness12", "awareness13",
                                        "insight4", "insight5", "insight14",
                                        "connection1", "connection17", "connection21",
                                        "connection2", "connection3", "connection11"),
                        nfactors = 3, fm = "ml", n.iter = 1000
)
print(fa_3factor)
print(fa_3factor$loadings, cut = .3)

fa_4factor <- psych::fa(data %>%
                          dplyr::select("purpose6.x", "purpose9.x", "purpose12.x", "purpose14.x",
                                        "awareness1.x", "awareness8.x", "awareness12.x", "awareness13.x",
                                        "insight4.x", "insight5.x", "insight14.x",
                                        "connection1.x", "connection17.x", "connection21.x",
                                        "connection2.x", "connection3.x", "connection11.x"),
                        nfactors = 4, fm = "ml", n.iter = 1000, cor="cor"
)
print(fa_4factor)
print(fa_4factor$loadings, cut = .3)
print(fa_4factor$residual)


fa_5factor <- psych::fa(hmindex_charlab_017 %>%
                          dplyr::select("purpose6", "purpose9", "purpose12", "purpose14",
                                        "awareness1", "awareness8", "awareness12", "awareness13",
                                        "insight4", "insight5", "insight14",
                                        "connection1", "connection17", "connection21",
                                        "connection2", "connection3", "connection11"),
                        nfactors = 5, fm = "ml", n.iter = 1000
)
print(fa_5factor)
print(fa_5factor$loadings,cut=.3)

```


## Convergent validity

```{r validity}

# Convergent validity with well-being scales
apaTables::apa.cor.table(dplyr::select(
  hmindex_charlab_017,awareness,connection,insight,purpose,hmindex,
  satwlife,who,psydistress,lonely,epconnect
),filename="/Users/tammirakral/Documents/HMIx/figures/wellbeing_corTable.doc",)

# Convergent validity for awareness comparison measures
apaTables::apa.cor.table(dplyr::select(
  hmindex_charlab_017,awareness,
  chimea_act, maas,chimea_awa, esq_attn
),filename="/Users/tammirakral/Documents/HMIx/figures/awareness_corTable.doc",)

# Convergent validity for connection comparison measures
apaTables::apa.cor.table(dplyr::select(
  hmindex_charlab_017,connection,
  gentrust, posrels, compassion
),filename="/Users/tammirakral/Documents/HMIx/figures/connection_corTable.doc",)

# Convergent validity for insight comparison measures
apaTables::apa.cor.table(dplyr::select(
  hmindex_charlab_017,insight,
  chimea_dec,chimea_rel,ders_lack_emoreg, ders_nonaccept,reappraisal
  ),filename="/Users/tammirakral/Documents/HMIx/figures/insight_corTable.doc")

# Convergent validity for purpose comparison measures
apaTables::apa.cor.table(dplyr::select(
  hmindex_charlab_017,purpose,
  francis_purpose,costin_purpose,meaning
  ),filename="/Users/tammirakral/Documents/HMIx/figures/purpose_corTable.doc")

```


## Inspect data

Quality checks & descriptives

```{r inspect}

# Descriptives for all (reformated) variables
varDescribe(data[9:51])

# Check descriptives filtered by subjects who have HMIdx data at T1 & T2
retest=data[!is.na(data$purpose6.x) & !is.na(data$purpose6.y),]
length(retest$AnonymizedStudentID)
varDescribe(retest[9:56])

### Quality assurance descriptives and distributions -----

# Check missingness across all variables
(map(
  data,
  ~ sum(is.na(.))
))

# Get demographics
length(which(data$gender == "female"))
length(which(data$gender == "male"))
length(which(data$gender == "nonbinary"))
length(which(data$gender == "other"))
length(which(is.na(data$gender)))

length(which(is.na(data$race)))
length(which(data$race == 1)) # native american/ aboriginal
length(which(data$race == 2)) # black
length(which(data$race == 3)) # east asian
length(which(data$race == 4)) # latino
length(which(data$race == 5)) # native hawaiian/ pacific islander
length(which(data$race == 6)) # other/ unknown
length(which(data$race == 7)) # south asian
length(which(data$race == 8)) # white

```

## Compute HMIdx scales

Filter for subjects with response times under 5 minutes & remove NAs

```{r compute}

T1insight_items <- data %>%
  dplyr::select(
    insight4.x, insight5.x, insight14.x
  )

T2insight_items <- data %>%
  dplyr::select(
    insight4.y, insight5.y, insight14.y
  )

psych::alpha(T1insight_items)
psych::alpha(T2insight_items)

# Awareness
T1awareness_items <- data %>%
  dplyr::select(
    awareness1.x,  awareness8.x, awareness12.x, awareness13.x
  )

T2awareness_items <- data %>%
  dplyr::select(
    awareness1.y,  awareness8.y,awareness12.y, awareness13.y
  )

psych::alpha(T1awareness_items)
psych::alpha(T2awareness_items)

# Connection
T1connection_items <- data %>%
  dplyr::select(
    connection1.x, connection2.x, connection3.x, connection11.x, connection17.x, connection21.x
  )

T2connection_items <- data %>%
  dplyr::select(
    connection1.y, connection2.y, connection3.y, connection11.y, connection17.y, connection21.y
  )

psych::alpha(T1connection_items)
psych::alpha(T2connection_items)

# scale computation: purpose ----
T1purpose_items <- data %>%
  dplyr::select(
    purpose6.x, purpose9.x, purpose12.x, purpose14.x
  )

T2purpose_items <- data %>%
  dplyr::select(
    purpose6.y, purpose9.y, purpose12.y, purpose14.y
  )

psych::alpha(T1purpose_items)
psych::alpha(T2purpose_items)

# whole hm index -----
hmidx_items <- data %>%
  dplyr::select(
    purpose6.x,
    purpose9.x,
    purpose12.x,
    purpose14.x, connection1.x, connection2.x, connection3.x, connection11.x, connection17.x, connection21.x, awareness1.x, awareness8.x, awareness12.x, awareness13.x, insight4.x, insight5.x, insight14.x
  )
psych::alpha(hmidx_items)

hmidx_items_retest <- data %>%
  dplyr::select(
    purpose6.y,
    purpose9.y,
    purpose12.y,
    purpose14.y, connection1.y, connection2.y, connection3.y, connection11.y, connection17.y, connection21.y, awareness1.y, awareness8.y, awareness12.y, awareness13.y, insight4.y, insight5.y, insight14.y
  )
psych::alpha(hmidx_items_retest)

```

## Calculate intra-class correlations (ICCs)

Verify correct model/ ICC version.

```{r icc}

data=data[!is.na(data$awareness.x),]

# select variable(s)
data$var1=data$hmindex.x
data$var2=data$hmindex.y
vars=c("var1","var2")
test=data[vars]

#test$var1=as.numeric(test$var1)
#test$var2=as.numeric(test$var2)

test2=test[!is.na(test$var1) & !is.na(test$var2),]
cor(test2)
# psych package ICC function can handle missingness

ver1ICC=ICC(test) # full data set
ver1ICC$results
# ICC(test2) # filtered for only non-missing data

# ICC for consistency (no missing values)
icc(test2,model = "twoway",
    type = "consistency",
    unit = "single",
    r0 = 0,
    conf.level = 0.95)

# ICC for agreement (no missing values)
icc(test2,model = "twoway",
    type = "agreement",
    unit = "single",
    r0 = 0,
    conf.level = 0.95)

    mretest=lm(var2~var1,data=test2)
    dof=df.residual(mretest)

        pY <- data.frame(var1=seq(min(test2$var1),max(test2$var1),length=dof))
        pY <- modelPredictions(mretest, pY) # predict values for the DVx

        scatterplot <- ggplot() + 
          geom_point(data=test2, aes(x = var1, y = var2),shape=1) + 
          geom_smooth(aes(ymin = CILo, ymax = CIHi,x = var1, y = Predicted), 
                      data = pY, stat = "identity", color="black") +
          theme_bw(base_size = 14) + 
          labs(y = 'Retest HM Index Total Score',
               x = 'Baseline HM Index Total Score')
        print (scatterplot)
        scatterplot

```
